---
title: "Social Network Analysis"
author: "Sander_MartinsGoncalves&Nisha_Antony"
date: "27/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Drawing a network using R
You are provided with a dataset (friendship011.csv), which contains the undirected network
for friendship among a group of students. You were also given a dataset with attributes for
each student. This includes:
- gender (1=female, 0=male),
- their grade (going from 0, the lowest to 10, the highest) and
- their top music preference (1=Rock Music, 2=Electronic Dance Music, 3=Pop Music, 4=Latin, 5=R&B)




```{r include=FALSE}
library("igraph")
library("equatiomatic")

frienship <- read.csv("../Dataset/Friendship011.csv", stringsAsFactors=FALSE, row.names=1)
profile_attribute <- read.csv("../Dataset/Attributes_Gender_Grade_Music.csv", stringsAsFactors=FALSE, row.names=1)

frienship <- as.matrix(frienship)

# Converting to matrix to igraph object
frienship_igraph <- graph_from_adjacency_matrix(frienship, mode = c("undirected"), diag = FALSE)

frienship_igraph # IGRPH UN-- 18 18 first val represents nodes and second val undirected edges

V(frienship_igraph)$name #name of the vertices
```

Checking if the order of the file is consistent
```{r}
sum(rownames(profile_attribute)!=colnames(frienship))
```
zero means that there is no different value regarding row/column. Good to go!

## 1.1. Draw the network using the Fruchterman-Reingold layout.

```{r echo=FALSE}
plot(frienship_igraph, layout=layout_with_fr, 
     vertex.size = 18,
     vertex.shape = "sphere",
     vertex.label.cex = 1,
     vertex.label.color = "grey3",
     vertex.label.dist = 1.5,
     vertex.label.degree=-pi/2)
```

## 1.2. Next draw the network again using the Fruchterman-Reingold layout but now with attributes gender, age and top music preference.

```{r echo=FALSE}
# Adding attributes to graph object
profile_attribute <- as.matrix(read.csv("../Dataset/Attributes_Gender_Grade_Music.csv", stringsAsFactors=FALSE, row.names=1))

# adding gender and grade to igraph obj
V(frienship_igraph)$Gender <- as.numeric(profile_attribute[, 1])
V(frienship_igraph)$Grade <- as.numeric(profile_attribute[, 2])
V(frienship_igraph)$Music <- as.numeric(profile_attribute[, 3])

frienship_igraph

# setting color, size and shape for attributes
V(frienship_igraph)$shape[V(frienship_igraph)$Gender == 0] <- "square"
V(frienship_igraph)$shape[V(frienship_igraph)$Gender == 1] <- "sphere"

V(frienship_igraph)$color[V(frienship_igraph)$Music == 1] <- "red"
V(frienship_igraph)$color[V(frienship_igraph)$Music == 2] <- "goldenrod1"
V(frienship_igraph)$color[V(frienship_igraph)$Music == 3] <- "green"
V(frienship_igraph)$color[V(frienship_igraph)$Music == 4] <- "black"
V(frienship_igraph)$color[V(frienship_igraph)$Music == 5] <- "olivedrab3"

plot(frienship_igraph, layout=layout_with_fr,
     vertex.size = V(frienship_igraph)$Grade*2,
     vertex.label.cex = 1,
     vertex.label.color = "grey3",
     vertex.label.dist = 1.5)

```


## 1.3. Create a legend for the attributes.

```{r echo=FALSE}
plot(frienship_igraph, layout=layout_with_fr,
     vertex.size = V(frienship_igraph)$Grade*2,
     vertex.label.cex = 1,
     vertex.label.color = "grey3",
     vertex.label.dist = 1.5,)

legend("bottomright", legend=c("Female", "Male"), 
       pch=21, pt.bg=c("black", "red"), pt.cex=2, bty="n")

legend("topleft", inset=.02,
   c("Rock Music", "Electronic Dance Music","Pop Music", "Latin", "R&B"),fill =frienship_igraph$Music , horiz=FALSE, cex=0.8)



# legend(x=.9, y=.9, legend=c("Female", "Male"), 
#        pch=21, pt.bg=c("black", "red"), pt.cex=2, bty="n")
```

## 1.4. Add the script of how you obtained the drawings.
Can be seen above.




# 2. Centrality

>Next, use the dataset (friendship011.csv) to calculate the centrality of all the nodes.

## 2.1. Calculate Freeman’s closeness centrality for each node and interpret your results.

> Provide the closeness centrality score for each node.

```{r echo=FALSE}
centrality <- closeness(frienship_igraph, normalized = T)
print("Centrality of nodes in descending order")
sort(centrality, decreasing = T)
```

### 2.1.1 What does it mean substantively in this case?

If we consider the flow of information, where we want to spread the information than the most efficient person to spread it would be "Jean" (information reach all network faster) and the last efficient one would be "Ellen" (takes long for the information to reach all network). In this case, the more central a person (node) is, the closer he/she is from all other people (nodes) in the network.


### 2.1.2 Who is most closeness central?

```{r echo=FALSE}

print("Highest centrality value: ")
centrality[which.max(centrality)]
```

### 2.1.3 Provide details of the meaning of the value.

A person that has the highest number in the network means that he/she is the most influence person on this network and it is defined by the number of links that this person has.


### 2.1.4 Make sure to discuss the reference points for the closeness measure. What is the minimum and maximum value?
```{r echo=FALSE}
print("Highest centrality: ")
print(centrality[which.max(centrality)])
print("Lowest centrality: ")
print(centrality[which.min(centrality)])
```
In the case where we have an information being spread by the most influencer person in the network(Jean), Ellen will be the last person (node) to receive it.

## 2.2. Calculate the betweenness centrality for each node.

> 2.2.1 Provide the betweenness centrality score for each node.

```{r echo=FALSE}
betweenness_centrality <-  sort(betweenness(frienship_igraph, directed = F, normalized = T), decreasing = T)
```

### 2.2.1 What does it mean substantively in this case?
It measures the control of information that a person have in the network (bridge concept). A person that is considered the shortest path to highest number of nodes has more control of information. 


### 2.2.2 Who is most betweenness central?

```{r betweenness centrality, echo=FALSE}
print("Highest betweenness centrality: ")
print(betweenness_centrality[which.max(betweenness_centrality)])
```

### 2.2.3 Provide details of the meaning of the value.

In some cases, certain parts of the network can only be reached through a certain node, that confers "power" to the node in a way that it could control the flow of information of the network

### 2.2.4 Make sure to discuss the reference points for the betweenness measure. What is the minimum and maximum value?

```{r echo=FALSE}
print("Maximum value:")
print(betweenness_centrality[which.max(betweenness_centrality)])
print("Maximum value:")
print(betweenness_centrality[which.min(betweenness_centrality)])
```

# 3. Density and reciprocity    
> Figure 1 provides the advice network among 9 employees in a team.

The figure 1 was created in order to do the calculations.
```{r include=FALSE}
a <- c(0, 1, 0,	0, 0,	0, 0,	0, 0)

b <- c(1,	0, 0,	0, 0,	0, 1,	0, 0)

c <- c(0,	1, 0,	0, 0,	0, 0,	0, 0)

d <- c(0,	0, 0,	0, 1, 0, 1,	0, 0)

e <- c(0,	0, 0,	0, 0,	0, 1,	0, 0)

f <- c(0,	0, 0,	0, 1,	0, 0,	0, 0)

g <- c(0, 1, 1, 1, 1, 1, 0, 1, 0)

h <- c(0, 1, 0, 0, 0, 0, 0, 0, 1)

i <- c(1, 0, 0, 0, 0, 0, 0, 0, 0)

emp_network <- data.frame(a, b, c, d, e, f, g, h, i)

x = c("a", "b", "c", "d", "e", "f", "g", "h", "i")
rownames(emp_network) <- x
emp_network <- as.matrix(emp_network)

# Density = number of ties/number of possible ties
density = 17/(9*8)
```


```{r Figure 1, echo=FALSE}
# Converting to matrix to igraph object
emp_igraph <- graph_from_adjacency_matrix(emp_network, mode = c("directed"))
plot(emp_igraph, layout=layout_with_fr, 
     vertex.size = 18,
     vertex.shape = "sphere",
     vertex.label.cex = 1,
     vertex.label.color = "grey3",
     vertex.label.dist = 1.5,
     vertex.label.degree=-pi/2)
edge_density(emp_igraph)
```




## 3.1. Calculate the density for this network and interpret.
Density = number of ties/number of possible ties
density = $\frac{17}{9*8}$
```{r echo=FALSE}
paste("density = ",17/(9*8))
```

> Make sure to show how you came to this answer.

### 3.1.1 What do you conclude substantively 
> (taking into account that the network is “Who did you ask advice from in the last week?”)?

There are more people asking advice (to receive) that than being asked (to give).


### 3.1.2 Make sure to discuss the reference points for the measure. What is the minimum and maximum value?

Maximum number of ties:
```{r}
paste("density of g = ",9/(9*8))
```

Minimum number of ties:
```{r}
paste("density of c ou f = ",2/(9*8))
```


# 3.2. Calculate the arc-based reciprocity index for this network and interpret.
> Make sure to show how you came to this answer.

> Arc-based Reciprocity formula:

> $\frac{r}{r+u}$ ${r = }$ reciprocity connections, ${u = }$ non reciprocity connections

```{r include=FALSE}
# Arc-based Reciprocity
reciprocity(emp_igraph)
r = 8;
u = 9;
arc_reci = function(r, u) {
  return(r / (r + u))
}
  

# Since "g" has max number of ties, we calculate the max val with respect to "g"

arc_rec_g = arc_reci(6, 3)
```


```{r echo=FALSE}
r = 8
u = 9 
print("If we consider the whole network")
paste(" reciprocity connections: ", r)
paste(" connections not reciprocity: ", u)
paste("Arc-based Reciprocity: ",arc_reci(r,u))


```

### 3.2.1 What do you conclude substantively?
> (taking into account that the network is “Who did you ask advice from in the last week?”)



### 3.2.2 Make sure to discuss the reference points for the measure. What is the minimum, expected and maximum value?

```{r}
arc_rec_a = arc_reci(, )
arc_rec_b = arc_reci(, )
arc_rec_c = arc_reci(, )
arc_rec_d = arc_reci(, )
arc_rec_e = arc_reci(, )
arc_rec_f = arc_reci(, )
arc_rec_g

```


# 4. Homophily/Heterophily for ego
Figure 2 shows the friendship network among students (with gender Male and Female).

## 4.1. Calculate the EI-index for person b in Figure 2.
> Make sure to show how you came to this answer.

EL-index = $\frac{(Ext - Int)}{(Ext + Int)}$ 

Ext = different type as node
Int = Same type as node

EL-index (b) = (2 - 2)/(2 + 2)
```{r echo=FALSE}
(2 - 2)/(2 + 2)
```


### 4.1.1 What do you conclude substantively?

> (taking into account that it is a friendship among students)

> Make sure to discuss the reference points for the measure.

EI score of 0 means that an equal number of alters are of both the same "type" as the ego, and different types. Here, Person-b is a female and has 2 male friends and 2 female friends. Equally quantity of nodes types.


## 4.2. Next, calculate Yule’s Q for person b in Figure 2.

> Make sure to show how you came to this answer.

$Q_i = \frac{ad - bc}{ad + bc}$

a = Nodes who are the same type as ego and Directly connected to ego
c = Nodes who are the same type as ego and *not* directly connected to ego

b = Nodes who are a different type than ego and directly connected to ego
d = Nodes who are a different type than ego and *not* directly connected to ego

```{r Yule's Q, include=FALSE}
qi = function(a,b,c,d){
  QI = (a*b-b*c)/(a*d+b*c)
  return(QI)
}
```

```{r}
qi(a = 2, b = 3, c = 2, d = 3)

```


### 4.2.1 What do you conclude substantively?

> (taking into account that it is a friendship among students)

> Make sure to discuss the reference points for the measure.




### 4.2.2 What is the difference with the EI index?

Yule’s Q varies from -1 (perfect heterophily) to +1 (perfect homophily) and a zero value would mean no pattern (i.e. equaly distributed in terms of types), same as the EI index.